---
title: "20160317_mass_spec_unfolding_analysis"
author: "Andrew Nguyen"
date: "2016-March-17"
output: pdf_document
---
#

#load library
```{r libraries}
library(ggplot2)
library(tidyr)
library(minpack.lm)
#functions

#fitting logistic curve 
#iwll help to plot if you just put parameters in
fud<-function(T=seq(25,50,1),Tm=40,slope=.5,max=1,min=0){
  y<-min+ (max-min)/(1+exp((-slope*(Tm-T))))
  return(y)
  }
plot(seq(25,50,1),fud(max=1,min=0),pch=15,col="black")
#nls function estimate the minimum unfolding



#nls fit
nls.fit<-function(data=data){
  y<-nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=data, start=list(slope=.5,Tm=40,min=.2), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000))
  return(coef(y))
  }
```

#load in data and visualize
```{r load and visualize}
p<-read.csv("../Data/20160317_parsed_pb17-10_unfolding_mass_spec_dat.csv")
str(p)

#getting rid of 25
p1<-p[,-5]
str(p1)

#temp points
Temp<-c(15,20,30,35,40,45,50,55,60)
#make a few plots
plot(Temp,p1[1,3:11],pch=16,cex=3,main="Unknown protein",ylab="Fraction non-denatured") # first row

plot(Temp,p1[38,3:11],pch=16,cex=3)#row 38
#info for row 38
p1[38,1:2]
test<-as.data.frame(t(rbind(Temp,p1[38,3:11])))
#test<-as.data.frame(t(rbind(Temp,p1[129,3:11]))) # no singular gradient
names(test)<-c("T","unfolding")

fit1<-nls(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=test, start=list(slope=.3,Tm=40,min=.2), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000))
summary(fit1)
plot(Temp,p1[38,3:11],pch=16,cex=3,ylim=c(0,1.3),main="ATP synthase subunit f, mitochondrial",ylab="Fraction non-denatured")
lines(seq(15,60,.1),fud(T=seq(15,60,.1),Tm=36.0947,slope=0.84522,min=0.31959),lwd=2,col="blue",lty="dotdash")

#tyring nlsLM

#no fit
test_fail<-as.data.frame(t(rbind(Temp,p1[1,3:11])))
#test_fail<-as.data.frame(t(rbind(Temp,p1[138,3:11])))
names(test_fail)<-c("T","unfolding")
#plot(Temp,p1[138,3:11],pch=16,cex=3)
plot(Temp,p1[1,3:11],pch=16,cex=3)

fit_fail<-nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=test_fail, start=list(slope=.3,Tm=35,min=.2), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000))
summary(fit_fail)

#should work
fit2<-nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=test, start=list(slope=.3,Tm=35,min=.2), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000))
summary(fit2)

# to see if the function worked
summary(fit_fail)$convInfo$stopMessage
summary(fit2)$convInfo$stopMessage
```
#looking for high tm proteins
```{r looking for high tm proteins}
test<-as.data.frame(t(rbind(Temp,p1[100,3:11])))
#test<-as.data.frame(t(rbind(Temp,p1[129,3:11]))) # no singular gradient
names(test)<-c("T","unfolding")

fitted<-nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=test, start=list(slope=.3,Tm=35,min=.2), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000))
summary(fitted)$parameters
```

#testing out data parsing
```{r testing out data parsing}
#making a dataset with aa gene that cant be fitted and one that can be
test$Gene<-rep("Gene1",length(test$T))
test_fail$Gene<-rep("Gene2",length(test_fail$T))
new_test_dat<-as.data.frame(rbind(test,test_fail))

##apply the function
play_func<-sapply(split(new_test_dat,list(new_test_dat$Gene)),function(x){ifelse(summary(nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=x, start=list(slope=.3,Tm=35,min=.2), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000)))$convInfo$stopMessage=="Relative error in the sum of squares is at most `ftol'.","worked","no")})

play_func
```

#lets do this for the real data set
workflow:
1) figure out which ones I can fit curves to
2) figure out how to grab their parameters (Tm, slope, min)
3) Graph and Esimate thermodynamics

#1)figure out which ones I can fit curves to
##changing wide to long format
```{r}
#before changing to wide to long format
head(p1)
p2<-tail(p1,150);dim(p2)

p2.long<-gather(p2,Description,unfolding,X15:X60)
p2.long.sort<-p2.long[order(p2.long[,2]),-6]

#adding temps
p2.long.sort$T<-rep(c(15,20,30,35,40,45,50,55,60),length(p2.long.sort$unfolding)/9)
str(p2.long.sort)
head(p2.long.sort,11)

p2.long.sort$Accession<-as.factor(p2.long.sort$Accession)
```
##fit models to all genes
```{r fit models to all genes}
library(plyr)
#cant fit to proteins that dont work, subset out the 60 degree temperature
all.fit<-sapply(split(p2.long.sort,list(p2.long.sort$Accession)),function(x){ifelse(summary(nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=x, start=list(slope=.5,Tm=37,min=.3), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000)))$convInfo$stopMessage=="Relative error in the sum of squares is at most `ftol'.","worked","no")})

all.fit

#param.fit
param.fit<-sapply(split(p2.long.sort,list(p2.long.sort$Accession)),function(x){summary(nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=x, start=list(slope=.5,Tm=37,min=.3), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000)))$parameters[,1]})
head(param.fit)
t(param.fit)
parameterss<-as.data.frame(cbind(t(param.fit),row.names(t(param.fit))))
parameterss
parameterss$slope<-as.numeric(as.character(parameterss$slope))
parameterss$Tm<-as.numeric(as.character(parameterss$Tm))
parameterss$min<-as.numeric(as.character(parameterss$min))

str(parameterss);dim(parameterss)
sub.param<-subset(parameterss,parameterss$Tm < 40 )
str(sub.param);dim(sub.param)

hist(sub.param$Tm)
hist(sub.param$slope)
hist(sub.param$min)

cor_param<-princomp(sub.param[,1:3],scale=T)
cor_param
cor_param$loadings[,1:3]

#write.csv(t(param.fit),"param_fits.csv")
#param.p.values
param.p<-sapply(split(p2.long.sort,list(p2.long.sort$Accession)),function(x){summary(nlsLM(unfolding ~ min+ (1-min)/(1+exp((-slope*(Tm-T)))),data=x, start=list(slope=.5,Tm=37,min=.3), trace=TRUE,control=nls.control(warnOnly = TRUE, tol = 1e-05, maxiter=1000)))$parameters[,4]})
head(param.p)
param_p<-as.data.frame(cbind(t(param.p),row.names(t(param.p))))
#write.csv(t(param.p),"param_p_values.csv")
```

#figures predictions of models
```{r predicting new values from model and plotting}
str(sub.param)
names(sub.param)[4]<-"Accession"
#ddply(sub.param,.(Accession),summarize,function(x) fud(T=seq(15,60,.1),Tm=x$Tm,slope=x$slope,min=x$min))
list_predictions<-sapply(split(sub.param,list(sub.param$Accession)),function(x) {fud(T=seq(15,60,.1),Tm=x$Tm,slope=x$slope,min=x$min)})

predi<-as.data.frame(do.call("rbind", list_predictions),stringAsFactors=FALSE)
predi$Accession<-sub.param$Accession
str(predi)

predi.long<-gather(predi,Accession,unfolding,V1:V451)
names(predi.long)[2]<-"Temp"
predi.long$Temp<-as.character(predi.long$Temp)
str(predi.long)
#library(stringer)
tmp<-unlist(strsplit(predi.long$Temp,split="V"))
length(na.omit(as.numeric(tmp)))
dim(predi.long)

predi.long$Tfac<-na.omit(as.numeric(tmp))
predi.long.sorted<-predi.long[order(predi.long$Accession),]
head(predi.long.sorted)

ddply(predi.long.sorted,.(Accession),summarize,k=length(Tfac))


predi.long.sorted$T<-rep(seq(15,60,.1),139)

#now i can plot the predicted values

ggplot(data=predi.long.sorted,aes(x=T,y=unfolding,colour=Accession))+geom_line()

```


#estimating the gibbs free energy
```{r estimating the gibbs free energy}
p1$Accession<-as.factor(p1$Accession)
str(p1)
names(sub.param)[4]<-"Accession"
dim(sub.param)
library(dplyr)
Kinetics_estimates<-inner_join(sub.param,p1,by=c("Accession"))
dim(Kinetics_estimates)

DeltaG_3<-function(Cp=5,T=(seq(-60,60,.1)+273.15),Tm=35+273.15,slope=.3){
  G<-(slope*Tm)*(1-(T/Tm))-(Cp*((Tm-T)+T*log(T/Tm)))
  m<-cbind(seq(-60,60,.1),G)

  return(max(m[,2]))

  }
#cbind(seq(-60,60,.1),DeltaG_3())
mo<-lm(Tm~calc..pI,data=Kinetics_estimates)
summary(mo)
plot(Kinetics_estimates$calc..pI,Kinetics_estimates$Tm)
abline(summary(mo))
#CP estimates
((Kinetics_estimates$X..AAs*12)/1000)*(pi/180)

#ddply(Kinetics_estimates,.(Accession),summarize,function(x) DeltaG_3(Cp=(x$X..AAs*12)/1000,slope=x$slope,Tm=Tm))
Gibbs<-sapply(split(Kinetics_estimates,list(Kinetics_estimates$Accession)),function(x){DeltaG_3(Cp=(x$X..AAs*12)/1000*(pi/180),slope=x$slope,Tm=x$Tm)})
hist(Gibbs,20,xlab="Gibbs free of unfolding")
range(Gibbs)
#Cp=(x$X..AAs*12)/1000*(pi/180),
```

